Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2012-06-29 3dimpvt/paper.tex
removing references to step 1

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data collected by
  modeling systems. Camera pose recovery in such systems often suffers
  from inaccuracies, resulting in visible discontinuties when images
  are projected adjacently onto a plane for texturing. We propose two
  approaches for reducing discontinuities in texture mapping 3D models
  made of planar surfaces. The first one is tile based and can be used
  for images and planes at arbitrary angles relative to each
  other. The second one, which we refer to as our seam minimization
  approach, results in a more seamless texture but is only applicable
  where camera axes for images are closely aligned with plane normals
  of the surfaces to be textured. The effectiveness of our approaches
  are demonstrated on two indoor datasets.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simulatenously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}.This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

@@

In this paper, we present a number of approaches to texture mapping 3D
models of indoor environments made of planar surfaces in the presence
of uncertainty and noise in camera poses. In particular, we consider a
human-operated backpack system with a number of laser range scanners
as well as 2 cameras facing left and right, each equipped with fisheye
lenses reaching an approximately 180$^{\circ}$ field of view and
taking photos at a rate of 5 Hz. With the cameras and laser scanners
active, the human operator wearing the backpack takes great care to
walk a path such that every wall in the desired indoor environment is
traversed and scanned lengthwise at least once, resulting in many
thousands of captured images.

Applying multiple localization and loop-closure algorithms on the data
collected by the ondoor sensors \cite{chen2010indoor,
  kua2012loopclosure, liu2010indoor}, the backpack is
localized\footnote{In this paper, we use the terms localization and
  pose recovery interchangeably, in that they both refer to recovering
  position and orientation.}  over its data collection period. This
requires recovering the 6 degrees of freedom for the backpack as well
as all its sensors, including the cameras rigidly mounted on it. Once
this is complete, the data from the laser range scanners is used to
generate a 3D point cloud of the surrounding environment and a 3D
planar model is generated \cite{sanchez2012point}. To do so, we
approximate normal vectors for each point in the point cloud by
gathering neighboring points within a small radius and processing them
through principal component analysis. These normal vectors allow for
the classification and grouping of adjacent points into structures
such as walls, ceilings, floors, and staircases. A RANSAC algorithm is
then employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model. This model, consisting of
multiple 2D polygonal planes in 3D space, along with the set of images
captured by the backpack's cameras and their noisy 3D poses, can be
considered the input to our texture mapping problem.

This problem straddles the fields of image-based localization
refinement as well as image alignment and mosaicing, and thus our
proposed solution shares many similarities with past research in those
areas, which will be examined in Section \ref{sec:existingApproaches}. However, given our knowledge of approximate camera poses as well as the 3D geometry of the environment to be textured, we can more accurately and efficiently texture our planes.

@@

In all subsequent sections, we discuss the process of texture mapping
a single plane, as the texturing of each of our planes is independent
and can be completed in parallel.

@@

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images with which we must texture the target
plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
the camera's intrinsic parameters, such as focal length and image
center, as well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and localization algorithms
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure} and are quite
noisy.

Because the backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each plane. Our goal in creating
a texture mapping process is to decide which of these images should be
used, and where their contents should map onto the plane, in order to
eliminate any visual discontinuities or seams that would suggest that
the plane's final texture is not composed of a single continuous image.

@@

We choose to work with rectangular units to ensure that borders
between any two distinct images in our final texture are either
horizontal or vertical. Since most environmental features inside
buildings are horizontal or vertical, any seams in our texture
intersect them minimally and are likely to be less noticeable.

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the $P_i$ camera matrices. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, the center
of $t$, and every other plane \cite{rayintersection}.

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while texture mapping a given tile. By using the
same image across tile boundaries, we can eliminate a discontinuity
altogether. If this is not possible because a tile is not visible in
images chosen by its neighbors, using similar images across a tile boundary will result in minimal discontinuities.

@@

If no satisfactory image is found in the first subset, we check the
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but were taken at a
similar location and time, suggesting that their localization and
projection are very similar. Again, if no viable image is found
according to the same criteria, we search the entire set of candidate
images, selecting based on the same scoring function from Figure
\ref{fig:scoringFunction}.

@@

Before examining these approaches, we first obtain a set of images to
work with.  Rather than perform camera or image adjustments across the
many thousands of images acquired in a typical data collection, we opt
to work with the more limited set of images corresponding to those
chosen by the direct mapping approach, without caching, of Section
\ref{sec:directMapping}. This set of images constitutes a reasonable
candidate set for generating a final seamless texture since it meets
three important criteria. First, each tile on the plane is covered by
at least one image in this set; this ensures no holes in the final
texture. Second, images are all selected according to the scoring
function in Figure \ref{fig:scoringFunction}, ensuring reasonable
camera distances and angles. Third, as a side result of the scoring
function, selected images are only viable for the tiles near their
center of projection. Thus, there should be plenty of overlap between
selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review approaches towards refining and
combining their projections.

One way to align image projections is to refine their noisy camera
poses, such that they become pixel accurate. This can be done by using
image correspondences to adjust each image's camera matrix over 6
degrees of freedom (DOF) \cite{coorg1997matching,
  liu2010indoor, shum2000systems, szeliski1997creating}. Using our same modeling datasets, recent work
followed this approach by solving a non-linear optimization problem to
refine camera poses during backpack localization \cite{liu2010indoor,
  chen2010indoor}. This approach however suffers from error
accumulation and drift as shown in Figure \ref{fig:mosaic3D}(a), as
well as high complexity and runtime.

Refining camera poses is analogous to image alignment or mosaicing,
which perform translations on images in order to align them directly.
One common approach to image alignment is to iteratively match images
on intensity differences in selected patches, often in a spatial
hierarchy \cite{hager1998efficient, lucas1981iterative,
  shum2000systems, szeliski2006image}. Another widely-used approach is
to perform feature matching, which requires the detection of visual
features and their existence in multiple images
\cite{brown2007automatic, cho2003automatic,
  mikolajczyk2005performance, szeliski2006image,
  yun2003comprehensive}. Notably, both approaches rely on the presence
of multiple visual references, as when texturing detailed objects at
high resolutions \cite{bernardinimultiplescans, wangmultipleviews}, or
when creating large scene panoramas \cite{agarwalapanoramas,
  bernardinimultiplescans, debevecviewdependent,
  snavelyphototourism}. In contrast, our indoor datasets have a high
prevalence of bare walls and ceilings, resulting in few reference
points that can be used to judge alignment.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homography, a pixel in the $nth$ image must be translated into the
first image's coordinates by multiplying by the $3\times3$ matrix $H_1
H_2 H_3 ... H_n$. Any error in one of these homography matrices is
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(b) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Even with many features spread across this plane,
the mosaicing produces errors that causes straight lines to appear as
waves on the plane, despite the fact that it was generated after
careful hand tuning. Many planes with fewer features simply failed
outright.

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{pipeline.pdf}
  \caption{Our proposed method for seamless texture mapping.}
  \label{fig:pipeline}
\end{figure}

@@

\subsection{Image Rotation}
\label{sec:projectionAndRotation}
We begin with the projection of all images onto the plane. This is
done in the same way as the approaches in Section
\ref{sec:simpleTextureMapping}. Rather than work with all 6 DOF, which is computationally intensive, we choose to work only in two dimensions, performing 2D rotations and translations on the image projections, as errors in either lead to the greatest discontinuities.

We perform rotations using Hough transforms, which
detect the presence and orientation of linear features in our
images. Rather than match the orientation of such features in each
image, we simply apply rotations such that the strongest near-vertical
features are made completely vertical. This is effective for vertical
planes in indoor models, since they usually consist of parallel
vertical lines corresponding to doors, wall panels, rectangular
frames, etc. If features in the environment are not vertical, e.g. for
floors and ceilings, or are not parallel to each other, this step is
skipped.

@@

SIFT matches determine $d^x$ and $d^y$ distances between each pair of
features for two images on the plane, though these distances may not
always be the same for different features.

Since indoor environments often contain repetitive features such as
floor tiles or doors, we need to ensure that SIFT-based distances are
reliable. In order to mitigate the effect of incorrect matches and
outliers, the RANSAC framework \cite{fischler1981random} is used for a
robust estimate of the optimal $d^x_{i,j}$ and $d^y_{i,j}$ distances
between two images $i$ and $j$. The RANSAC framework handles the
consensus-building machinery, and requires a fitting function and a
distance function. For this application, the fitting function simply
finds the average distance between matches in a pair of images. The
distance function for a pair of points is chosen to be the difference
between those points' SIFT match distance and the average distance
computed by the fitting function. We use a 10 pixel outlier threshold,
so that SIFT matches are labeled as outliers if their horizontal or
vertical distances are not within 10 pixels of the average distance
computed by the fitting function.

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates that we generated a SIFT-based distance between
images 1 and 2, images 1 and 3, images 2 and 3, etc.

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there are often cases where there is a break in
the chain of images, meaning that no SIFT matches are found between
one segment of the plane and another. In this case we add rows to the
$A$ matrix and observations to the $\vec{\gamma}$ vector that contain
the original noisy $x$ and $y$ distance estimates generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor}. Another
way to do this is to add rows for all neighboring pairs of images and
solve a weighted least squares problem where the SIFT distances are
given a higher weight i.e. 1, and the noisy distances generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor} are given
a smaller weight i.e. 0.01.

@@

As mentioned earlier, the mobile backpack system used to capture data
in this paper contains 2 cameras with $180^\circ$ fisheye lenses
facing to the right and left. Since most human operators do not carry
the backpack in a perfectly upright position and are bent forwards at
15 to 20 degrees with respect to the vertical direction, the
undistorted fisheye images are head on with respect to vertical walls,
but at an angle with respect to horizontal floors and ceilings. This
is depicted in Figure \ref{fig:projectionAngles}. These oblique camera
angles translate into textures that span extremely large areas once
projected. Using the tile-based texture mapping criteria from Figure
\ref{fig:scoringFunction}, such projections have high or low scores
depending on the location on the plane, as shown in Figure
\ref{fig:projectionAngles}(b). The tiling approach in Section
\ref{sec:simpleTextureMapping} is thus a good choice for texturing
floors and ceilings, as it uses the parts of image projections that
are viable choices for their respective plane locations, e.g. areas
near point A in Figure \ref{fig:projectionAngles}(b), but not near
point B.

For wall planes however, most images are taken from close distances
and more or less head-on angles, resulting in higher resolution
fronto-parallel projections. As a result, for each tile on a wall
plane, the cost function of Figure \ref{fig:scoringFunction} is
relatively flat with respect to a large number of captured images, as
they are all more or less head on. Thus it is conceivable to use a
different texturing strategy for walls so as to minimize the visible
seams for texture mapped planes. One way to do so is to choose the
smallest possible set of images that (a) covers the entire plane and
(b) minimizes the amount of visible seams. A straigntforward cost
function that accomplishes the latter is the Sum of Squared pixel
Differences (SSD) in overlapping regions between all pairs of images,
after they have been aligned as described in Section
\ref{sec:projectionAndRotation}. Minimizing this cost function
encourages image boundaries to occur either in featureless areas, such
as bare walls, or in areas where images match extremely well.

@@

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a long hallway. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between images that overlap. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal cost $C_0$ to(from) all images that
meet the left(right) border of the plane.

We now solve the shortest path problem from the start node to the end
node. This provides a set of images completely covering the plane
horizontally, while minimizing the cost of seams between images.

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. To address this, we solve for
a new shortest path after modifying the edges chosen by the original
shortest path solution such that they have higher cost than all other
edges in our DAG, and solve for a new shortest path. This ensures that
these edges are not chosen again unless they are the only available
option. Our second shortest path solution results in a new set of
chosen images, of which we only retain those that cover areas not
covered by the first set. This process can be repeated as many times
as needed, until holes are no longer present. This method is not as
optimal as a 2D-coverage solution would be, but it is a fast
approximation, and adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the posters in the images, which have clear misalignment over seams in Figure \ref{fig:compareAll}(c), but are much more aligned in Figure \ref{fig:compareAll}(d). This is because the seam minimization approach
directly reduces the cost of each image boundary, while the tile
caching method uses a scoring function that only approximates this
effect. Furthermore, seam minimization guarantees the best selection
of images, while the sequential tile caching method may select images
early on that turn out to be poor choices once subsequent tiles have
been processed. The seam minimization approach is also far less
intensive in terms of memory usage and runtime, as it does not require
discretizing planes or images.

@@

\subsection{Blending}
\label{sec:blending}
We now apply the same blending process to the two texturing methods:
localization refinement followed by either tile caching or seam
minimization.

@@

In this paper, we have developed an approach to texture map models
with noisy camera localization data. We are able to refine image
locations based on feature matching, and robustly handle outliers. The
tile-based mapping approach can be used to texture both simple
rectangular walls as well as complex floor and ceiling geometry. We
also presented a seam minimization texturing method that produces
seamless textures on vertical planes where multiple head-on images are
available. Each of these approaches is highly modular, and easily
tunable for different environments and acquisition hardware.

@@

  \centering \subfloat[][]{\includegraphics[width=3in]{fullmodel.png}}
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) a ceiling, (c) an entire floor, (d) a full model}
  \label{fig:results}
\end{figure}

@@

\appendix
\section*{Appendix A. Occlusion Masking}
\label{sec:occlusionMasking}
For the seam minimization process, it is important that images contain
only content that should be mapped onto the target plane in
question. The tiling approach checks occlusion for each tile as it is
being textured, whereas the seam minimization approach uses entire
images.  Occlusion checks thus need to be performed over the entirety
of each image to determine available areas for texture mapping.

Fortunately, by virtue of our indoor environments, the vast majority
of surface geometry is either horizontal or vertical, with high
amounts of right angles. This means that after masking out occluded
areas, image projections will remain largely rectangular. We can be
efficient by recursively splitting each image into rectangular pieces,
and performing the same occlusion checks used in the tiling process
where needed. To occlude out rectangular pieces, their texture is
removed, as untextured areas are never used for texture mapping.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data collected by
  modeling systems. Camera pose recovery in such systems often suffers
  from inaccuracies, resulting in visible discontinuties when
  successive images are projected adjacently onto a plane for
  texturing. We propose two approaches to reduce discontinuities in
  texture mapping 3D models made of planar surfaces. The first one is
  tile based and can be used for images and planes at arbitrary angles
  relative to each other. The second one results in a more seamless
  texture, but is only applicable where camera axes for images are
  closely aligned with plane normals of the surfaces to be
  textured. The effectiveness of our approaches are demonstrated on
  two indoor datasets.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simulatenously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

@@

In this paper, we present a number of approaches to texture mapping 3D
models of indoor environments made of planar surfaces in the presence
of uncertainty and noise in camera poses. In particular, we consider
data obtained from a human-operated backpack system with a number of
laser range scanners as well as 2 cameras facing left and right, each
equipped with fisheye lenses reaching an approximately 180$^{\circ}$
field of view and taking photos at a rate of 5 Hz. Applying multiple
localization and loop-closure algorithms on the raw data collected by
the onboard sensors \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, the backpack is localized\footnote{In this paper, we
  use the terms localization and pose recovery interchangeably, in
  that they both refer to recovering position and orientation.}  over
its data collection period. This requires recovering the 6 degrees of
freedom for the backpack as well as the cameras rigidly mounted on
it. Once this is complete, the data from the laser range scanners is
used to generate a 3D point cloud of the surrounding environment, from which a
3D planar model is created \cite{sanchez2012point}. This model, consisting of 2D
polygonal planes in 3D space, along with the set of images captured by
the backpack's cameras and their noisy 3D poses, can be considered the
input to our texture mapping problem.

This problem straddles the fields of image-based localization
refinement as well as image alignment and mosaicing, and thus our
proposed solution shares many similarities with past research in those
areas, which will be examined in Section
\ref{sec:existingApproaches}. However, we can exploit known
approximate camera poses as well as the 3D geometry of the environment
to more accurately and efficiently texture the planes.

@@

In all subsequent sections, we discuss the process of texture mapping
a single plane, as the texturing of each plane is independent and can
be completed in parallel.

@@

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images to texture the target
plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
the camera's intrinsic parameters, such as focal length and image
center, as well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and localization algorithms
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure} and are quite
noisy.

Because the backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each plane. Our goal in creating
a texture mapping process is to decide which of these images should be
used, and where their contents should map onto the texture, in order
to eliminate any visual discontinuities or seams that would suggest
that the plane's final texture is not composed of a single continuous
image.

@@

We choose to work with rectangular units to ensure that borders
between any two distinct images in the final texture are either
horizontal or vertical. Since most environmental features inside
buildings are horizontal or vertical, any seams in our texture
intersect them minimally and are likely to be less noticeable.

In order to select an image for texturing a tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the $P_i$ camera matrices. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, the center
of $t$, and every other plane \cite{rayintersection}.

@@

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while texture mapping a given tile. By using the
same image across tile boundaries, we can eliminate a discontinuity
altogether. If this is not possible because a tile is not visible in
images chosen by its neighbors, using similar images across tile
boundaries also leads to less noticeable discontinuities.

@@

If no satisfactory image is found in the first subset, we check the
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but are taken at a
similar location and time, suggesting that their localization and
projection are quite similar. Again, if no viable image is found
according to the same criteria, we search the entire set of candidate
images, selecting based on the same scoring function from Figure
\ref{fig:scoringFunction}.

@@

Image projections can be aligned by refining their noisy camera poses,
using image correspondences to adjust each image's camera matrix over
6 degrees of freedom (DOF) \cite{coorg1997matching, liu2010indoor,
  shum2000systems, szeliski1997creating}. Applying such approaches to
our datasets however results in error accumulation and drift as shown
in Figure \ref{fig:mosaic3D}(a), as well as high complexity and
runtime \cite{liu2010indoor}.

Another common approach to image alignment is to
iteratively match images on intensity differences in selected patches,
often in a spatial hierarchy \cite{hager1998efficient,
  lucas1981iterative, shum2000systems, szeliski2006image}. Another
widely-used approach is to perform feature matching, which requires
the detection of visual features and their existence in multiple
images \cite{brown2007automatic, cho2003automatic,
  mikolajczyk2005performance, szeliski2006image,
  yun2003comprehensive}. Notably, both approaches rely on the presence
of multiple visual references, as when texturing detailed objects at
high resolutions \cite{bernardinimultiplescans, wangmultipleviews}, or
when creating large scene panoramas \cite{agarwalapanoramas,
  bernardinimultiplescans, debevecviewdependent,
  snavelyphototourism}. In contrast, our indoor datasets have a high
prevalence of bare walls and ceilings, resulting in few reference
points that can be used to judge alignment.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homography, a pixel in the $nth$ image must be translated into the
first image's coordinates by multiplying by the $3\times3$ matrix $H_1
H_2 H_3 ... H_n$. Any error in one of these homography matrices is
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(b) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Even with many features spread across this plane,
the mosaicing produces errors that causes straight lines to appear as
waves on the plane, despite the fact that it is generated after
careful hand tuning. Many planes with fewer features simply failed
outright using this approach.

@@

\begin{figure}
  \centering
  \includegraphics[width=3.5in]{pipeline.pdf}
  \caption{Our proposed method for seamless texture mapping.}
  \label{fig:pipeline}
\end{figure}

@@

\subsection{Image Rotation}
\label{sec:projectionAndRotation}
We begin with the projection of all images onto the plane. This is
done in the same way as in Section
\ref{sec:simpleTextureMapping}. Rather than work with all 6 DOF, which
is computationally intensive, we choose to work only in two
dimensions, performing 2D rotations and translations on the image
projections, as errors in either lead to the greatest discontinuities.

We perform rotations using Hough transforms, which detect the presence
and orientation of linear features in our images. Rather than match
the orientation of such features in each image, we simply apply
rotations such that the strongest near-vertical features are made
completely vertical. This is effective for vertical planes in indoor
models, since they usually consist of parallel vertical lines
corresponding to doors, wall panels, rectangular frames, etc. If
features in the area are not vertical, e.g. for floors and
ceilings, this step is skipped.

@@

SIFT matches determine $d^x$ and $d^y$ distances between each pair of
features for two images on the plane, though these distances may not
always be the same for different features. Since indoor environments
often contain repetitive features such as floor tiles or doors, we
need to ensure that SIFT-based distances are reliable. In order to
mitigate the effect of incorrect matches and outliers, the RANSAC
framework \cite{fischler1981random} is used for a robust estimate of
the optimal $d^x_{i,j}$ and $d^y_{i,j}$ distances between two images
$i$ and $j$. The RANSAC framework handles the consensus-building
machinery, and requires a fitting function and a distance
function. For this application, the fitting function simply finds the
average distance between matches in a pair of images. The distance
function for a pair of points is chosen to be the difference between
those points' SIFT match distance and the average distance computed by
the fitting function. We use a 10 pixel outlier threshold, so that
SIFT matches are labeled as outliers if their horizontal or vertical
distances are not within 10 pixels of the average distance computed by
the fitting function.

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there is often a break in
the chain of images, meaning that no SIFT matches are found between
one segment of the plane and another. In this case we add rows to the
$A$ matrix and observations to the $\vec{\gamma}$ vector that contain
the original noisy $x$ and $y$ distance estimates from the original
camera poses. Another way to do this is to add rows for all
neighboring pairs of images and solve a weighted least squares problem
where the SIFT distances are given a higher weight i.e. 1, and the
noisy distances generated by the original camera poses
\cite{chen2010indoor, liu2010indoor} are given a smaller weight
e.g. 0.01.

@@

As mentioned earlier, the mobile backpack system used to capture data
in this paper contains 2 cameras with $180^\circ$ fisheye lenses
facing to the right and left. Since most human operators do not carry
the backpack in a perfectly upright position and are bent forwards at
15 to 20 degrees with respect to the vertical direction, the
undistorted fisheye images are head on with respect to vertical walls,
but at an angle with respect to horizontal floors and ceilings. This
is depicted for wall and ceiling planes in Figures
\ref{fig:projectionAngles}(a) and \ref{fig:projectionAngles}(b)
respectively. These oblique camera angles for ceilings translate into
textures that span extremely large areas once projected, as shown in
Figure \ref{fig:projectionAngles}(b). Using the tile-based texture
mapping criteria from Figure \ref{fig:scoringFunction}, such
projections have highly varying scores depending on the location on
the plane, as shown in Figure \ref{fig:projectionAngles}(b). Thus, the
tiling approach in Section \ref{sec:simpleTextureMapping} is a reasonable
choice for texturing floors and ceilings, as it uses the parts of
image projections that are viable choices for their respective plane
locations, e.g. areas near point A in Figure
\ref{fig:projectionAngles}(b), but not near point B.

For wall planes however, most images are taken from close distances
and more or less head-on angles, resulting in higher resolution
fronto-parallel projections. As a result, for each tile on a wall
plane, the scoring function of Figure \ref{fig:scoringFunction} is
relatively flat with respect to a large number of captured images, as
they are all more or less head on. Thus, the scoring function is less
signifcant for walls, and it is conceivable to use a different
texturing strategy for them so as to minimize visible seams in texture
mapped planes. This can be done by choosing the smallest possible set
of images that (a) covers the entire plane and (b) minimizes the
amount of visible seams between them. A straigntforward cost function
that accomplishes the latter is the Sum of Squared Differences (SSD)
of pixels in overlapping regions between all pairs of images, after
they have been rotated as described in Section
\ref{sec:projectionAndRotation}. Minimizing this cost function
encourages image boundaries to occur either in featureless areas, such
as bare walls, or in areas where images match extremely well.

@@

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a long hallway. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal cost $C_0$ to(from) all images that
meet the left(right) border of the plane.

We now solve the shortest path problem from the start node to the end
node. This results in a set of images completely covering the plane
horizontally, while minimizing the cost of seams between images.

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. To address this, we solve for
a new shortest path after modifying the edges chosen by the original
shortest path solution such that they have higher cost than all other
edges in the DAG, and solve for a new shortest path. This ensures that
these edges are not chosen again unless they are the only available
option. Our second shortest path solution results in a new set of
chosen images, of which we only retain those that cover areas not
covered by the first set. This process can be repeated as many times
as needed, until holes are no longer present. This method is not as
optimal as a 2D-coverage solution would be, but it is a fast
approximation, and adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the
posters in the images, which have clear misalignment over seams in
Figure \ref{fig:compareAll}(c), but are much more aligned in Figure
\ref{fig:compareAll}(d). This is because the seam minimization
approach directly reduces the cost of each image boundary, while the
tile caching method uses a scoring function that only approximates
this effect. Furthermore, seam minimization guarantees the best
selection of images, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. The seam minimization approach
is also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images\footnote{ For the seam minimization
  process, it is important that images only contain content that
  should be mapped onto the target plane in question. The tiling
  approach checks occlusion for each tile as it is being textured,
  whereas the seam minimization approach uses entire images.
  Occlusion checks thus need to be performed over the entirety of each
  image to determine available areas for texture mapping a given
  plane.  }.

@@

\subsection{Blending}
\label{sec:blending}
We now apply the same blending process to the two texturing methods:
image alignment followed by either tile caching or seam
minimization.

@@

In this paper, we have developed an approach to texture map models
with noisy camera localization data. We are able to refine image
locations based on feature matching, and robustly handle outliers. The
tile-based mapping approach can be used to texture both simple
rectangular walls as well as complex floor and ceiling geometry. We
also presented a seam minimization texturing method that produces
seamless textures on planes where multiple head-on images are
available. Each of these approaches is highly modular, and easily
tunable for different environments and acquisition hardware.

@@

  \centering \subfloat[][]{\includegraphics[width=3in]{fullmodel.png}}
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) a ceiling, (c) an entire floor, (d) a full model.}
  \label{fig:results}
\end{figure}

@@
